---
title: "EVI-Pro Lite Load Profiles"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 5)
options(warning.length = 2000L)
```

Load libraries and functions. The openEVI() function sources all functions except randVector(), so all we need are these two.

```{r}
library(future.apply)
library(ggplot2)
library(data.table)
library(magrittr)
library(foreach)
library(doFuture)

source("functions/func_openEVI.R")
source("functions/func_randVector.R")

theme_set(theme_bw() + theme(strip.background = element_blank()))
```

Specify user variable options stored in a vector. Where there are different permutations of user variable sets, store as a list of vectors. future.apply will parallel iterate through the different permutations. 

```{r}
#REMOVE THIS WHEN READY. KEEP FOR TROUBLE SHOOTING AT THE MOMENT
#mean_vmt <- 40 #mean daily vehicle miles traveled

#temp_list <- c(0, 10, 15, 20) #temp in deg C. Options are -20, -15, -10, -5, 0, 5, 10, 15, 20, 25, 30, 35, 40

# weights_list <- list("rural" = list(weights = NULL,
#                                     temp = 15,
#                                     mean_vmt = 50,
#                                     loc_class = "rural"),
#                      "urban" = list(weights = NULL,
#                                     temp = 25,
#                                     mean_vmt = 30,
#                                     loc_class = "urban"))
```

Create set of test vectors
```{r}
fleet_size_vec <- list(500)
mean_dvmt_vec <- c(15)
pev_type_vec <- list("BEV100" = c(0, 0, .30, .70))
veh_class_vec <- list("Sed80_Suv20" = c(0.8, 0.2))
loc_class_vec <- list("urban")
home_power_vec <- list("HA100_L2_20" = c(0.80, 0.20, 0))
work_power_vec <- list("L2_80" = c(0, 0.8))
pref_vec <- list("Home100" = c(1.0, 0.0))

# I want temp to be length > 1
temp_vec <- c("-20C", # celsius
               "-10C",
               "0C",
               "10C",
               "20C",
               "30C",
               "40C") 
```

Full vectors
```{r}
# count of total number of PEVs
fleet_size_vec <- list(500,
																							1000,
																							2500,
																							5000,
																							10000,
																							20000,
																							50000) 

# average daily miles traveled per vehicle
mean_dvmt_vec <- c(15,
                   20,
                   25,
                   35,
                   45,
                   50) 

# Temperature
temp_vec <- c("-20C", # celsius
               "-10C",
               "0C",
               "10C",
               "20C",
               "30C",
               "40C"
               ) 

# PEV Type (confirm the meaning of the labels) (def named wrong)
pev_type_vec <- list("BEV100" = c(0, 0, .30, .70), # PHEV20, PHEV50,BEV100, BEV250; all BEV
                     "BEV_DOM" = c(.10, .15, .25, .50), # BEV Dominant
                     "PHEV_DOM" = c(.25, .50, .10, .15), # PHEV Dominant
                     "EQUAL_DIST" = c(.15, .35, .15, .35)) # PHEV/BEV Equal share

# adding the vehicle class variable
veh_class_vec <- list("Sed80_Suv20" = c(0.8, 0.2), # sedan dominant
																						"Sed50_Suv50" = c(0.5, 0.5), # equal distribution
																						"Sed20_Suv80" = c(0.2, 0.8)) # suv dominant

# Geography
loc_class_vec <- list("urban")

# Day of Week (included in code)
# c(Weekday, Weekend)

# Home Access and Power 																# L1, L2, % without access to home power 
home_power_vec <- list("HA100_L2_20" = c(0.80, 0.20, 0), #100% access to home power, 80% of those L1
                       "HA100_L2_80" = c(0.20, 0.80, 0), #100% access to home power, 80% of those L2
                       "HA100_L2_50" = c(0.5, 0.5, 0), #100% access to home power, 50% of those L2
                       "HA75_L2_20" = c(0.6, 0.15, 0.25), #75% access to home power, 80% of those L1
                       "HA75_L2_80" = c(0.15, 0.6, 0.25), #75% access to home power, 80% of those L2
                       "HA75_L2_50" = c(0.375, 0.375, 0.25), #75% access to home power, 50% of those L2
                       "HA50_L2_20" = c(0.4, 0.1, 0.5), #50% access to home power, 80% of those L1
                       "HA50_L2_80" = c(0.1, 0.4, 0.5), #50% access to home power, 80% of those L2
                       "HA50_L2_50" = c(0.25, 0.25, 0.5)) #50% access to home power, 50% of those L2

# drop all L2 cases, replace with 50/50 split
# make these an 80/20 split

# Work power
work_power_vec <- list(
	"L2_80" = c(0, 0.8),
	"L2_50" = c(0.5, 0.5),
	"L2_20" = c(0.8, 0.2)
	)
	
	# Home work preference
	pref_vec <- list(
	"Home100" = c(1.0, 0.0),
	"Home80" = c(0.8, 0.2),
	"Home60" = c(0.6, 0.4)
	)
```

Permutations calculation
```{r}
length(fleet_size_vec) * 
	length(mean_dvmt_vec) * 
	length(pev_type_vec) * 
	length(veh_class_vec) * 
	length(work_power_vec) *
	2 * # for day of week
	length(home_power_vec) *
	length(pref_vec)
	length(temp_vec)
```



Run the function that wraps the whole thing on each element in the list.

```{r message=FALSE}
Sys.time()
registerDoFuture() #allows foreach() to use the future package when the %dopar% operator is called
plan(multiprocess, workers = 4, gc = TRUE) #set future plan, make sure at least 1 core is reserved, call garbage collector
options(future.globals.maxSize = 25000 * 1024^2) #Set max RAM allowed for global variables to 10GB

lapply(temp_vec, function(temp) {
	


#for(i in 1:length(temp_vec)) {

  # load the raw data for sedans
  # evi_data_path_sedan <- paste0("../../data/preprocessed/evi_raw/", temp_vec[i], ".rds")
  # evi_raw <- readRDS(evi_data_path_sedan)
  evi_raw <- readRDS(paste0("../../data/preprocessed/evi_raw/", temp, ".rds"))
  # add column for vehicle class
  evi_raw[, "vehicle_class" := "Sedan"]
  
  # Repeat for SUVs
  # evi_data_path_suv <- paste0("../../data/preprocessed/evi_raw_suv/", temp_vec[i], ".rds")
  # evi_raw_suv <- readRDS(evi_data_path_suv)
		evi_raw_suv <- readRDS(paste0("../../data/preprocessed/evi_raw_suv/", temp, ".rds"))
  # add column for vehicle class
  evi_raw_suv[, "vehicle_class" := "SUV"]
  
  # add one hundred million to the unique vids to distinguish from sedan data
  evi_raw_suv[, unique_vid := unique_vid + 100000000]
  
  # bind together into a single data.table
  evi_raw <- rbind(evi_raw_suv,
  																	evi_raw)
  
  rm(evi_raw_suv)
    # load the load profiles
  # loadprof_data_path_sedan <- paste0("../../data/preprocessed/load_profile/", temp_vec[i], ".rds")
  # evi_load_profiles <- readRDS(loadprof_data_path_sedan)
  evi_load_profiles <- readRDS(paste0("../../data/preprocessed/load_profile/", temp, ".rds"))
  
  # add vehicle class
  evi_load_profiles[, "vehicle_class" := "sedan"]
  
  # Repeat for SUVs
  # loadprof_data_path_suv <- paste0("../../data/preprocessed/load_profile/", temp_vec[i], ".rds")
  # evi_load_profiles_suv <- readRDS(loadprof_data_path_suv)
  evi_load_profiles_suv <- readRDS(paste0("../../data/preprocessed/load_profile/", temp, ".rds"))
  # add vehicle class
  evi_load_profiles_suv[, "vehicle_class" := "SUV"]
  
  # add one hundred million to unique vid to distinguish SUV class of vehicle
  evi_load_profiles_suv[, unique_vid := unique_vid + 100000000]
  
  # bind together into a single data.table
  evi_load_profiles <- rbind(evi_load_profiles_suv,
  																											evi_load_profiles)
  
  rm(evi_load_profiles_suv)
  
  gc()
  #Parallel sessions for permutations of all other variables
  fleet_load <-
      foreach(numveh = fleet_size_vec, .combine='rbind') %do% {
        sub_fleet_load <-
          foreach(pev = names(pev_type_vec), .combine='rbind', .options.future=list(scheduling=1)) %:%
          foreach(dvmt = mean_dvmt_vec, .combine='rbind') %:%
          foreach(pref = names(pref_vec), .combine='rbind') %:%
          foreach(home = names(home_power_vec), .combine='rbind') %:%
          foreach(work = names(work_power_vec), .combine='rbind') %:%
          foreach(loc = loc_class_vec, .combine='rbind') %:%
        		foreach(vclass = names(veh_class_vec), .combine = "rbind") %dopar% {
          	
            #Generate fleet load profile for a single permutation of fleet, pev, dvmt, pref, home, work
            load_to_bind <- openEVI(evi_raw,
            																								evi_load_profiles,
            																								numveh,
                                    pev_type_vec[[pev]],
                                    dvmt,
                                    pref_vec[[pref]],
                                    home_power_vec[[home]],
                                    work_power_vec[[work]],
                                    loc,
            																								veh_class_vec[[vclass]])
       
            # add columns to indicate variable permutations of this run
            load_to_bind[, ':='(loc_class = loc,
                                temp_c = temp,
                                fleet_size = numveh,
                                mean_dvmt = dvmt,
                                pev_dist = pev,
                                pref_dist = pref,
                                home_access = strsplit(home,c("_"))[[1]][1],
                                home_power_dist = strsplit(home,c("_"))[[1]][2],
                                work_power_dist = work,
            																				vehicle_class_dist = vclass)]
            
            # Reduce fleet activity down to 24-hour aggregated load profile
            # Note: Remove this step, and reduce the number of variable permutations, to confirm imposed weight distributions, vehicle counts, etc.
            load_to_bind[time_of_day > 24,time_of_day := time_of_day - 24]
            load_to_bind <- load_to_bind[, .(kw = sum(avg_kw)),
                                         by = c("loc_class",
                                                "temp_c",
                                                "fleet_size",
                                                "mean_dvmt",
                                                "pev_dist",
                                                "pref_dist",
                                                "home_access",
                                                "home_power_dist",
                                                "work_power_dist",
                                                "day_of_week",
                                                "pev_type", 
                                                "dest_type", 
                                                "dest_chg_level",
                                                "time_of_day",
                                         							"vehicle_class_dist")]
            setkey(load_to_bind,
            							loc_class,
            							temp_c,
            							fleet_size,
            							mean_dvmt,
            							pev_dist,
            							pref_dist,
            							home_access,
            							home_power_dist,
            							work_power_dist,
            							day_of_week,
            							pev_type,
            							dest_type,
            							dest_chg_level,
            							time_of_day)
            
            gc()
            
            return(load_to_bind)
        }
      }
  
# If I want to drop columns or change temp to integer it would be here
 fleet_load[, "loc_class" := NULL]
 
#change temp, fleet size, dvmt to integer
 fleet_load$temp_c  <-
	as.integer(gsub("C", "", fleet_load$temp_c))

# convert other columns to integer
fleet_load[, fleet_size := as.integer(fleet_size)]
fleet_load[, mean_dvmt := as.integer(mean_dvmt)]
 
  
  #Write fleet_load out to disk
  fwrite(fleet_load,
  										file = paste0("outputs/testing/",
  																								gsub("-", "", Sys.Date()), "_", # date the run
  																								temp,
  																								"dev",
  																								".csv"))
  
  rm(evi_raw)
  rm(evi_load_profiles)
  rm(fleet_load)
  gc()
  
  # adding a check on the memory usage at each iteration for testing
  print(object.size(x=lapply(ls(), get)), units="Mb")
#} # end of for loop
})
Sys.time()
```

Spot check the output

```{r}
# remove residual data usage
gc(full = T)

# get files (currently only one file for testing)
out_path <- dir("outputs/testing", full.names = T)

# load data
check <- fread(out_path)

# subset as the full data is too large for a ggplot
sub_check <- check[fleet_size == 50000 &
                     mean_dvmt == 50 &
                     pev_dist == "BEV100" &
                     pref_dist == "Home50" &
                     home_access == "HA50" &
                     home_power_dist == "L225",
                   .(avg_kw = sum(kw)),
                   by=c("day_of_week",
                        "pev_type",
                        "dest_type",
                        "dest_chg_level",
                        "time_of_day",
                   					"vehicle_class_dist")]

test1 <- ggplot(data = sub_check, aes(x = time_of_day, y = avg_kw)) +
  geom_area(aes(fill = dest_chg_level)) +
  facet_grid(vehicle_class_dist ~ dest_type)

test1
```

```{r}
library(dplyr)

sub_check <-
	check %>% 
			filter(fleet_size == 5000 & 
											mean_dvmt == 50 &
											pev_dist == "BEV25" & # (pev_distribution) labels currently wrong
											pref_dist == "Home50" &
											home_access == 'HA50' &
											day_of_week == "weekday" &
											pev_type == "BEV250") %>% 
			group_by(day_of_week,
												pev_type,
												dest_type,
												dest_chg_level,
												time_of_day,
												vehicle_class_dist) %>% 
			summarise(avg_kw = sum(kw))


sub_check %>% 
	ggplot() +
	geom_area(aes(x =time_of_day, y = avg_kw, fill = dest_chg_level)) +
	facet_grid(vehicle_class_dist ~ dest_type) +
	labs(subtitle = unique(sub_check$pev_type))
```

```{r}
test2 <- ggplot(sub_check[dest_type == "Public" & day_of_week == "weekend" & pev_type == "BEV250",sum(avg_kw),by=c("time_of_day","dest_chg_level")], aes(x=time_of_day, y=V1)) +
  geom_line() +
  facet_grid(~dest_chg_level)

test3 <- ggplot(sub_check[,sum(avg_kw),by=c("day_of_week","dest_chg_level","time_of_day")], aes(x=time_of_day, y=V1)) +
  geom_area(aes(fill=dest_chg_level)) +
  facet_grid(~day_of_week)
```


```{r}
# test_dt <- rbindlist(flt_list)
# test_dt[,':='(loc_class = factor(loc_class, levels = loc_class_vec),
#               temp_c = factor(temp_c, levels = temp_vec),
#               pev_dist = factor(pev_dist, levels = names(pev_type_vec)),
#               pref_dist = factor(pref_dist, levels = names(pref_vec)),
#               home_power_dist = factor(home_power_dist, levels = names(home_power_vec)),
#               work_power_dist = factor(work_power_dist, levels = names(work_power_vec)))]

prof_24_by_class <- copy(test_dt)
prof_24_by_class[time_of_day > 24,time_of_day := time_of_day - 24]
prof_24_by_class <- prof_24_by_class[, .(kw = sum(avg_kw)),
                                by = c("loc_class",
                                       "temp_c",
                                       "fleet_size",
                                       "mean_dvmt",
                                       "pev_dist",
                                       "pref_dist",
                                       "home_power_dist",
                                       "work_power_dist",
                                       "day_of_week",
                                       "pev_type", 
                                       "dest_type", 
                                       "dest_chg_level",
                                       "time_of_day")]
setkey(prof_24_by_class,loc_class,temp_c,fleet_size,mean_dvmt,pev_dist,pref_dist,home_power_dist,work_power_dist,day_of_week,pev_type,dest_type,dest_chg_level,time_of_day)
write.csv(prof_24_by_class[temp_c == "-20C",.SD], file="output/Example_load-profile-structure_-20C.csv")

```

Check out the PEV type differences between scenarios.

```{r}
test_dt[temp_c == "-10C",.(pev_type = unique(pev_type),scenario = unique(loc_class),day_of_week = unique(day_of_week)),by=c("fleet_id","pev_dist")][, .N, by = c("pev_dist","pev_type", "scenario", "day_of_week")] %>%
  ggplot(aes(pev_type, N, fill = scenario)) + 
  geom_col(position = "dodge") + 
  facet_grid(pev_dist ~ day_of_week) +
  labs(title = "Number of Vehicles by type")

test_dt[temp_c == "-10C",.(vmt = unique(schedule_vmt),scenario = unique(loc_class),day_of_week = unique(day_of_week)),by=c("fleet_id","pev_dist")][, .N, by = c("pev_dist","vmt", "scenario", "day_of_week")] %>%
  ggplot(aes(vmt)) + 
  geom_histogram(binwidth = 10) + 
  facet_wrap(pev_dist ~ day_of_week) +
  labs(title = "Number of Vehicles by type")
```

Plot the 24 hour profile in a few different configurations.

```{r}
prof_24_by_class <- copy(test_dt)
prof_24_by_class[time_of_day > 24,time_of_day := time_of_day - 24]
prof_24_by_class <- prof_24_by_class[, .(kw = sum(avg_kw)),
                                by = c("loc_class",
                                       "temp_c",
                                       "fleet_size",
                                       "mean_dvmt",
                                       "pev_dist",
                                       "pref_dist",
                                       "home_power_dist",
                                       "work_power_dist",
                                       "day_of_week",
                                       "pev_type", 
                                       "dest_type", 
                                       "dest_chg_level",
                                       "time_of_day")]

prof_24_by_class[, .(avg_kw = sum(kw)), 
                        by = c("time_of_day",
                               "dest_chg_level",
                               "day_of_week")] %>%
  ggplot(aes(x = time_of_day)) +
  geom_area(aes(y = kw, fill = dest_chg_level)) +
  scale_fill_brewer("Charger Level", palette = "Dark2") +
  labs(x = "Time of day", y = "Average kw") + 
  facet_wrap(day_of_week ~ scenario, ncol = 2)

test <- prof_24_by_class[loc_class == "urban" &
                             temp_c == "-20C" &
                             fleet_size == 500 &
                             mean_dvmt == 15 &
                             pev_dist == "PHEV" &
                             pref_dist == "Home" &
                             home_power_dist == "HA100L1" &
                             work_power_dist == "AllL2" &
                             day_of_week == "weekday" &
                             pev_type == "PHEV20" &
                             dest_type == "Home" &
                             dest_chg_level == "L1",
                         kw, by=time_of_day]

fleet_load[, .(avg_kw = sum(kw)), 
                        by = c(
                                       "home_power_dist",
                                       "pev_dist",
                                       "pref_dist",
                                       "day_of_week",
                                       "time_of_day")] %>%
ggplot(aes(x = time_of_day)) +
  geom_area(aes(y = avg_kw, fill = dest_chg_level)) +
  scale_fill_brewer("Charger Level", palette = "Dark2") +
  labs(x = "Time of day", y = "Average kw") +
  facet_grid(day_of_week + temp_c ~ pev_dist + dest_type)

fleet_load[home_power_dist == "L225", .(avg_kw = sum(kw)), 
                        by = c(
                                       "home_access",
                                       "pev_dist",
                                       "pref_dist",
                                       "dest_chg_level",
                                       "day_of_week",
                                       "time_of_day")] %>%
ggplot(aes(x = time_of_day)) +
  geom_area(aes(y = avg_kw, fill = dest_chg_level)) +
  scale_fill_brewer("Charger Level", palette = "Dark2") +
  labs(x = "Time of day", y = "Average kw") +
  facet_grid(day_of_week + home_access ~ pev_dist + pref_dist)
```

How do the ecdfs compare?

```{r}
ggplot(prof_24_by_class, aes(avg_kw_by_class, color = dest_chg_level, linetype = scenario)) +
  stat_ecdf() +
  scale_color_brewer("Charger Level", palette = "Dark2") +
  scale_linetype("Location") +
  facet_grid(day_of_week ~ dest_chg_level)
```

Output data for NREL

```{r}
nrel_dt_1_hires <- flt_list$rural$load_profile[,.(temp_c,scenario,fleet_avg_vmt,day_of_week,pev_type,dest_type,dest_chg_level,fleet_id,session_id,time_of_day,avg_kw)]
setkey(nrel_dt_1_hires,day_of_week,fleet_id,session_id,time_of_day)
write.csv(nrel_dt_1_hires,file=paste0(getwd(),"/output/SEIN-Load-Profile-Example1-HiRes.csv"))

nrel_dt_1_lowres <- nrel_dt_1_hires[,.(avg_kw = sum(avg_kw)),by=c("temp_c","scenario","fleet_avg_vmt","day_of_week","pev_type","dest_type","dest_chg_level","time_of_day")]
setkey(nrel_dt_1_lowres,time_of_day)
write.csv(nrel_dt_1_lowres,file=paste0(getwd(),"/output/SEIN-Load-Profile-Example1-LowRes.csv"))

nrel_dt_2_hires <- flt_list$urban$load_profile[,.(temp_c,scenario,fleet_avg_vmt,day_of_week,pev_type,dest_type,dest_chg_level,fleet_id,session_id,time_of_day,avg_kw)]
setkey(nrel_dt_2_hires,day_of_week,fleet_id,session_id,time_of_day)
write.csv(nrel_dt_2_hires,file=paste0(getwd(),"/output/SEIN-Load-Profile-Example2-HiRes.csv"))

nrel_dt_2_lowres <- nrel_dt_2_hires[,.(avg_kw = sum(avg_kw)),by=c("temp_c","scenario","fleet_avg_vmt","day_of_week","pev_type","dest_type","dest_chg_level","time_of_day")]
setkey(nrel_dt_2_lowres,time_of_day)
write.csv(nrel_dt_2_lowres,file=paste0(getwd(),"/output/SEIN-Load-Profile-Example2-LowRes.csv"))
```

